PanLex-RDF

### 2013-01-01 paper/2012/SWJ-PanLex/paper.tex
Some more work

\title{PanLex}
\runningtitle{PanLex}
\todo{find a real title}

@@

\author[A]{\fnms{Patrick} \snm{Westphal}},
\author[A]{\fnms{Claus} \snm{Stadler}},
%\runningauthor{M. Martin et al.}
\address[A]{University of Leipzig, Institute of Computer Science, AKSW, Postfach 100920, D-04009 Leipzig, Germany\\
E-mail: \{lastname\}@informatik.uni-leipzig.de}
\todo{the e-mail address pattern won't work for both of us}

@@

Since Lexvo covers all ISO 639-3 language codes, interlinking could be done with
the Sparqlify-ML view in Figure~\ref{fig:ex:sparqlify-ml}.
For DBpedia we were interested in creating \emph{valid} and
thus \emph{dereferenceable} links.
Therefore, we iterated the \emph{titles} datasets\footnote{\url{http://wiki.dbpedia.org/Downloads38} - Note that more languages are available in the raw download section at \url{http://downloads.dbpedia.org/3.8/}},
which map non-localized DBpedia URIs to their page titles in the respective language.
For each language version we normalized the labels: 
Unicode NFKD\footnote{\url{http://unicode.org/reports/tr15/}} normalization and
removal of punctuation characters were applied.
% PanLex now uses NFC, not NFKD, for character normalization and does not remove punctuation.
The DBpedia resources were then mapped to the PanLex expression that was equal to the resource's normalized label in the respective language.
Table ~\ref{fig:plx-dbp-link-counts} summarizes the number of links obtained.
\begin{table}
\centering
\begin{scriptsize}
\begin{tabular}{llll}
\toprule
Language & 639-1 & 639-3 & Links \\
\midrule
English    & en & eng & 1,415,241 \\
German     & de & deu &   224,146 \\
French     & fr & fra &   187,364 \\
Italian    & it & ita &   147,485 \\
Spanish    & sp & spa &   117,056 \\
Portuguese & pt & por &   112,266 \\
Polish     & pl & pol &   110,974 \\
Russian    & ru & rus &    68,040 \\
Czech      & cs & ces &    28,767 \\
Catalan    & ca & cat &    27,779 \\
Korean     & ko & kor &    24,912 \\
Turkish    & tr & tur &    22,258 \\
Bulgarian  & bg & bul &    19,431 \\
Hungarian  & hu & hun &    18,203 \\
Redo Slovene & sl & slv &  4,636 \\
Greek      & el & ell &  1,112 \\
\bottomrule
\end{tabular}
\end{scriptsize}
\caption{Number of DBpedia links per language}
\label{fig:plx-dbp-link-counts}
\end{table}

@@

\section{Publishing}
\label{sec:publishing}
PanLex itself already provides direct access to the PostgreSQL database for scientists,
several user interfaces\footnote{\url{http://panlex.org/try/}} and an
API\footnote{\url{http://panlex.org/tech/doc/api/}}.
%The usage of these interfaces is not restricted in order to make the provided lexical
%translations as accessible as possible.
With our RDF conversion work, we complement these interfaces with
the Linked Data\footnote{\url{http://panlex.org/plx/TODO}} and the SPARQL
endpoint\footnote{\url{http://panlex.org/sparql} and \url{http://panlex.org/snorql}}.
An overview is shown in ~\autoref{fig:panlex-architecture}.
Our code for interlinking is
hosted on GitHub\footnote{\url{https://github.com/AKSW/PanLex-2-RDF}}.
\todo{Where to host the linksets?}

@@

\section{Conclusions and Future Work}
\label{sec:conclusion}
In this paper we describe the PanLex database and its conversion to RDF.
Based on our URI and vocabulary design, we created appropriate view definitions
for the virtual RDB to RDF solution \emph{Sparqlify} which carries out
the actual RDF transformation.
Furthermore, we interlinked the languages in PanLex with lexvo,
and created a total of \todo {number} derefenceable
links to DBpedia for expressions in 16 languages.
%We detailed the developed vocabulary and

In this work we also identified some weaknesses in
the orginal database design,
which we intend to overcome in the future:
The orginal dataset currently does not cleanly model information sources and approvers
as distinct entities.\todo{We should still map it like that with Sparqlify - resuse ap for the source}
Also, while it is possible to determine each meaning's and each denotation's approver, an approver can have multiple users with editorial privileges, so it is not possible to track each contributed, modified, and deleted meaning and denotation by editor.
Retaining this information seems beneficial for translation approches,
as this enables for example the attribution of qualities and relevances to particular editors.

>>>>

\title{PanLex-RDF}
\runningtitle{Countering language attrition with the Web of Data}
%\todo{find a real title}

@@

\author[A]{\fnms{Patrick} \snm{Westphal}},
\author[B]{\fnms{Claus} \snm{Stadler}},
\author[C]{\fnms{Jonathan} \snm{Pool}},

%\runningauthor{M. Martin et al.}

\address[A]{University of Leipzig, Institute of Computer Science, AKSW, Postfach 100920, D-04009 Leipzig, Germany\\
E-mail: \{lastname\}@informatik.uni-leipzig.de}
\address[B]{University of Leipzig}
\address[C]{TODO}
%\todo{the e-mail address pattern won't work for both of us}

@@

Since Lexvo covers all ISO 639-3 language codes, interlinking could be done with
the Sparqlify-ML view in Figure~\ref{fig:ex:sparqlify-ml}.
For DBpedia we were interested in creating \emph{valid} and
thus \emph{dereferenceable} links.
Therefore, we iterated the \emph{titles} datasets\footnote{\url{http://wiki.dbpedia.org/Downloads38} - Note that more languages are available in the raw download section at \url{http://downloads.dbpedia.org/3.8/}},
which map (non-localized) DBpedia URIs to their page titles in the respective language.
For each language version we normalized the labels: 
Unicode NFKD\footnote{\url{http://unicode.org/reports/tr15/}} normalization and
removal of punctuation characters were applied.
% PanLex now uses NFC, not NFKD, for character normalization and does not remove punctuation.
The DBpedia resources were then mapped to the PanLex expression that was equal to the resource's normalized label in the respective language.
Table ~\ref{fig:plx-dbp-link-counts} summarizes the number of links obtained.
\begin{table}
\centering
\begin{scriptsize}
\begin{tabular}{llll}
\toprule
Language & 639-1 & 639-3 & Links \\
\midrule
English    & en & eng & 1,415,241 \\
German     & de & deu &   224,146 \\
French     & fr & fra &   187,364 \\
Italian    & it & ita &   147,485 \\
Spanish    & sp & spa &   117,056 \\
Portuguese & pt & por &   112,266 \\
Polish     & pl & pol &   110,974 \\
Russian    & ru & rus &    68,040 \\
Czech      & cs & ces &    28,767 \\
Catalan    & ca & cat &    27,779 \\
Korean     & ko & kor &    24,912 \\
Turkish    & tr & tur &    22,258 \\
Bulgarian  & bg & bul &    19,431 \\
Hungarian  & hu & hun &    18,203 \\
Slovene    & sl & slv &    11,981 \\
Greek      & el & ell &     1,112 \\
\midrule
Total      &    &     & 2,537,015 \\
\bottomrule
\end{tabular}
\end{scriptsize}
\caption{Number of DBpedia links per language}
\label{fig:plx-dbp-link-counts}
\end{table}

@@

\section{Publishing}
\label{sec:publishing}
PanLex itself already provides direct access to the PostgreSQL database for scientists,
several user interfaces\footnote{\url{http://panlex.org/try/}} and an
API\footnote{\url{http://panlex.org/tech/doc/api/}}.
%The usage of these interfaces is not restricted in order to make the provided lexical
%translations as accessible as possible.
With our RDF conversion work, we complement these interfaces with
the Linked Data
%\footnote{\url{http://panlex.org/plx/TODO}}
and the SPARQL
endpoint\footnote{\url{http://panlex.org/sparql} and \url{http://panlex.org/snorql}}.
An overview is shown in ~\autoref{fig:panlex-architecture}.
Our code for interlinking is
hosted on GitHub\footnote{\url{https://github.com/AKSW/PanLex-2-RDF}}.
\todo{Where to host the linksets?}

@@

\section{Conclusions and Future Work}
\label{sec:conclusion}
In this paper we describe the PanLex database and its conversion to RDF.
Based on our URI and vocabulary design, we created appropriate view definitions
for the virtual RDB to RDF solution \emph{Sparqlify} which carries out
the actual RDF transformation.
Furthermore, we interlinked the languages in PanLex with lexvo,
and created about 2.5 million links to DBpedia for expressions in 16 languages.
%We detailed the developed vocabulary and

In this work we also identified some weaknesses in
the orginal database design,
which we intend to overcome in the future:
The orginal dataset currently does not cleanly model information sources and approvers
as distinct entities.
Also, while it is possible to determine each meaning's and each denotation's approver, an approver can have multiple users with editorial privileges, so it is not possible to track each contributed, modified, and deleted meaning and denotation by editor.
Retaining this information seems beneficial for translation approches,
as this enables for example the attribution of qualities and relevances to particular editors.
